### 2022-03-07 paper/localpsf.tex
paper edits

Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation. The details of the interpolation will be described in Section \ref{sec:approximate_kernel_entries}, but the key principle behind the interpolation is the local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
%Specifically, we make the assumption that
%\begin{equation}
%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
%\end{equation}
%if $x_i$ is close to $x$. This is shown in Figure \ref{fig:mean_displacement_invariance}. Motivation for this assumption is discussed in Section \ref{eq:local_mean_displacement_invariance}. 
We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$. The weights in the linear combination are determined via preconditioned radial basis function interpolation.

The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H$, of $\Phi$, using the conventional adaptive cross H-matrix construction method. In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}. We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$. Schematically the overall H-matrix approximation proceeds as follows:
\begin{equation*}
\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
\end{equation*}
One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion. WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A. These H-matrix methods are fast and scale well to large problems.

@@

Our method of estimating the support of the functions $\convkernel_i$ was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}. In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\convkernel_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$. If the width of $\convkernel_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size. In \cite{RESOLUTION}, the width of $\convkernel_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion. In this paper, rather than probing $\Aop^T$ with random noise functions, we use specific constant, linear, and quadratic functions. Our method estimates the support of $\convkernel_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.

Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid \cite{PRODCONVFFT}. The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator \cite{PRODCONVKRYLOV}. However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform. Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations. Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}. Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we convert the product-convolution approximation to H-matrix format.

@@

Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}. In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators. Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product interpolation, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.

\section{Method}
\label{sec:method}
In this section we describe how we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$ (Section \ref{sec:moments}), the greedy ellipsoid packing procedure we use to choose batches of sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute impulse responses, $\phi_{x_i}$, in batches (Section \ref{sec:get_impulse_response}), how we interpolate the $\phi_{x_i}$ to approximate integral kernel entries $\Phi(y,x)$ (Section \ref{sec:approximate_kernel_entries}), finite element discretization of $\Phi$ (Section \ref{sec:finite_element_kernel}), and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$. We also describe how to (optionally) modify $\mathbf{A}_h$ to force it to be symmetric positive semi-definite. The complete algorithm is shown in Algorithm \ref{alg:construct_Atilde}.

@@

			Compute $\combresponse_k$ by applying $\Aop$ to the dirac comb for $\pointbatch_k$ (Section \ref{sec:get_impulse_response})

@@

\section{Impulse response moments}
\label{sec:moments}

@@

\subsection{Computational cost}

>>>>

Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation. The details of the interpolation will be described in Section \ref{sec:approximate_kernel_entries}, but the key principle behind the interpolation is the local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
%Specifically, we make the assumption that
%\begin{equation}
%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
%\end{equation}
%if $x_i$ is close to $x$. This is shown in Figure \ref{fig:mean_displacement_invariance}. Motivation for this assumption is discussed in Section \ref{eq:local_mean_displacement_invariance}. 
We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$. The weights are determined by preconditioned radial basis function interpolation.

The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H$, of $\Phi$, using the conventional adaptive cross H-matrix construction method. In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}. We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$. Schematically the overall H-matrix approximation proceeds as follows:
\begin{equation*}
\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
\end{equation*}
One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion. These H-matrix methods are fast and scale well to large problems. WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A.

@@

Our method of estimating the support of the functions $\convkernel_i$ was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}. In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\convkernel_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$. If the width of $\convkernel_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size. In \cite{RESOLUTION}, the width of $\convkernel_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion. In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions. Our method estimates the support of $\convkernel_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.

Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid \cite{PRODCONVFFT}. The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator \cite{PRODCONVKRYLOV}. However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform. Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations. Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}. Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted product-convolution approximations into H-matrix format.

@@

Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}. In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators. Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product approximation (a product-convolution approximation of the transpose of the operator).
%, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.

\section{Method}
\label{sec:method}
In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$. This section is organized as follows:
\begin{description}
	\item[Section \ref{eq:mean_and_covariance_estimation}:] We describe our ellipsoid estimate for the support of $\phi_x$.
	\item[Section \ref{sec:moments}:] We show how to compute $V,\mu,\Sigma$.
	\item[Section \ref{sec:sample_point_selection}:] We present the greedy ellipsoid packing procedure used to choose batches, $\pointbatch_b$, of sample points $x_i$.
	\item[Section \ref{sec:get_impulse_response}:] We describe how to compute $\phi_{x_i}$'s, in batches, $\eta_b$.
	\item[Section \ref{sec:approximate_kernel_entries}:] We show how we interpolate $\phi_{x_i}$'s to approximate $\Phi(y,x)$.
	\item[Section \ref{sec:finite_element_kernel}:] We discuss finite element discretization of $\Phi$.
	\item[Section \ref{sec:h_matrix_construction_short}:] We describe how we construct the H-matrices $\mathbf{\Phi}_H$ and $\mathbf{A}_H$.
	\item[Section \ref{sec:overall_cost}:] We review the total computational cost to construct $\mathbf{A}_H$, and list the cost to perform further linear algebra tasks involving $\mathbf{A}_H$ using H-matrix methods.
\end{description}
The optional method for making $\mathbf{A}_H$ positive semi-definite is described in Appendix \ref{sec:make_spd}. The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}.

%we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$ (Section \ref{sec:moments}), the greedy ellipsoid packing procedure we use to choose batches of sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute impulse responses, $\phi_{x_i}$, in batches (Section \ref{sec:get_impulse_response}), how we interpolate the $\phi_{x_i}$ to approximate integral kernel entries $\Phi(y,x)$ (Section \ref{sec:approximate_kernel_entries}), finite element discretization of $\Phi$ (Section \ref{sec:finite_element_kernel}), and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$. We also describe how to (optionally) modify $\mathbf{A}_H$ to force it to be symmetric positive semi-definite. 
%The complete algorithm is shown in Algorithm \ref{alg:construct_Atilde}.

@@

			Compute $\combresponse_k$ by applying $\Aop$ to the Dirac comb for $\pointbatch_k$ (Section \ref{sec:get_impulse_response})

@@

\subsection{Impulse response moments}
\label{sec:moments}

@@

\subsection{Computational cost}
\label{sec:overall_cost}
